---
title: "Motion clouds rivalry: Target detection - Target Eccentricity"
author: "Egor Ananyev"
date: "August 25, 2017"
output:
    html_document:
        toc: true
---

<style type="text/css">
body, td{
  font-family: Cambria;
}
code.r{
  font-family: Inconsolata;
}
</style>

# Preparing the data

## Loading the packages and data directories
```{r set-options, message=F}
options(width=108)
out=F
# The list of subjects, the order of conditions, and the thresholds are derived from Subjects.xlsx
library(xlsx)
library(ggplot2)
library(plyr)
library(reshape)
library(matrixStats)
library(gridExtra)
library(lme4)
library(BayesFactor)
#library(splines)
db <- '/home/egor/Dropbox/' # on Linux
db <- '/Users/Egor/Dropbox/' # Windows
# db <- '~/Dropbox/' # on Mac
```

## Plot variables
```{r, message=FALSE, warning=FALSE}
# theme for plotting:
alpha <- .7
themefy <- function(p) {
    p <- p + theme_bw() + 
         theme(panel.grid.minor.x=element_blank(), panel.grid.minor.y=element_blank(),
            axis.text=element_text(size=8), axis.title=element_text(size=9),
            legend.text=element_text(size=8), legend.title=element_text(size=9),
            legend.key = element_blank(), legend.margin=margin(t=-.04, unit='in'),
            legend.background = element_rect(fill='transparent'),
            plot.title=element_text(face='bold'))
}
cc <- c('#e31a1c','#fdbf6f','#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99')
```

```{r}
xLab <- 'Target Peak Time (s)'
yLab <- 'Log Contrast Threshold'
yLab2 <- 'Normalized Contrast Threshold'
colLab <- expression(paste('\nTarget\nEccentricity (', degree, ')', sep=''))
# colLab2 <- expression(paste('Target Eccentricity (', degree, ')', sep=''))
# colLabType <- 'Mask Type'
yLim <- c(-1.4,-0.6)
yLimNorm <- c(-1.2,-0.7)
dodge <- position_dodge(width=0.1)
sumFn <- function(ss, subjStr='subj', xStr='targTpeak', grpStr='targEcc'){
    sumSubj <- ddply(ss, c(subjStr, xStr, grpStr), summarise,
                     mnS=mean(threshMean), se=sd(threshMean)/sqrt(length(threshMean))) #, .drop=F)
    sumSubjMn <- ddply(ss, c(subjStr), summarise, mnStot=mean(threshMean)) # total mean across conditions per subj
    sumSubj <- merge(sumSubj, sumSubjMn)
    sumSubj$normS <- - sumSubj$mnS / sumSubj$mnStot # normalized subject mean
    sumSubj$seNorm <- NA
    # sumSubj$mnS[is.na(sumSubj$mnS)] <- 0
    sumGrp <- ddply(sumSubj, c(xStr, grpStr), summarise,
                  mn=mean(mnS), se=sd(mnS)/sqrt(length(mnS)),
                  norm=mean(normS), seNorm=sd(normS)/sqrt(length(normS)))
    sumGrp$subj <- 'average'
    sumSubj <- rename(sumSubj, c(mnS='mn',normS='norm'))
    sumComb <- rbind(sumGrp, subset(sumSubj, select=-mnStot))
    sumComb$se[is.na(sumComb$se)] <- 0
    sumComb
}
plotAve <- function(pss, subjStr='subj', xStr='targTpeak', grpStr='targEcc', 
                    xlab=xLab, ylab=yLab, collab=colLab, yStr='mn', seStr='se'){
    pss$yMin <- pss[,yStr] - pss[,seStr]
    pss$yMax <- pss[,yStr] + pss[,seStr]
    pss[,grpStr] <- factor(pss[,grpStr])
    p <- ggplot(pss, aes_string(x=xStr, y=yStr, colour=grpStr, group=grpStr,
                            ymin='yMin', ymax='yMax')) +
        geom_point(position=dodge, size=1, alpha=alpha) + geom_line(position=dodge, alpha=alpha) +
        scale_x_continuous(breaks=c(.5,1,1.5), labels=c('0.5','1','1.5')) +
        geom_linerange(position=dodge, show.legend=F, alpha=alpha) +
        labs(x=xlab, y=ylab, colour=collab) + #ylim(0,1) + 
        guides(colour=guide_legend(keyheight=.3, default.unit='inch'))
    p <- themefy(p)
}
plotIndiv <- function(pss, subjStr='subj', xStr='targTpeak', grpStr='targEcc', 
                    xlab=xLab, ylab=yLab, collab=colLab, yStr='mn', seStr='se'){
    pss$yMin <- pss[,yStr] - pss[,seStr]
    pss$yMax <- pss[,yStr] + pss[,seStr]
    pss[,grpStr] <- factor(pss[,grpStr])
    p <- ggplot(pss, aes_string(x=xStr, y=yStr, colour=grpStr, group=grpStr,
                            ymin='yMin', ymax='yMax')) + 
        facet_wrap( ~ subj, ncol=4) +
        geom_point(position=dodge, size=1, alpha=alpha) + geom_line(position=dodge, alpha=alpha) +
        geom_linerange(position=dodge, show.legend=F, alpha=alpha) +
        scale_x_continuous(breaks=c(.5,1,1.5), labels=c('0.5','1','1.5')) +
        # scale_y_continuous(breaks=c(0,.25,.5,.75,1), labels=c('0','','0.5','','1'), limits=c(0,1)) +
        labs(x=xlab, y=ylab, colour=collab) + 
        guides(colour=guide_legend(keyheight=.3, default.unit='inch'))
    p <- themefy(p)
}
```

## Scaling function
```{r}
cent <- function(v){
    v <- apply(v,2,function(x){
        x <- x - mean(unique(x))
        x <- x / max(x)
    })
}
```

## Loading the data
```{r}
allDataDir <- paste(db,'Projects/mc/data_bv3/mcBv3_maskV',sep='')
dataDirs <- dir(allDataDir)
dataDirs <- dataDirs[grep('targEcc',dataDirs)]
colsOfInt <- c('participant', 'dom', 'session', 'mcBv', 'targTpeak', 'targXoff2', 'stairStart', 
               'meanRev6')
df <- data.frame()
for(curDir in dataDirs){
    print(curDir)
    curDf <- read.csv(paste(allDataDir,'/',curDir,'/',curDir,'.csv',sep=''))
    curDf <- curDf[,colsOfInt]
    df <- rbind(df, curDf[,colsOfInt])
}
ds <- rename(df, c(participant='subj', session='sess', meanRev6='thresh',
                   mcBv='maskV', targXoff2='targEcc'))
ds$targEcc <- round(ds$targEcc / 35,1)
ds$maskType <- 'dynamic'
# ds$maskV <- factor(ds$maskV)
ds$maskV <- round(ds$maskV * 60 / 35, 1)
ds$maskV[ds$maskV<0.05] <- 0
ds$maskType[ds$maskV<=0.01] <- 'static'
head(ds)
```

## Transformed data sets

### By target eccentricity
```{r}
thresh <- ddply(ds, .(subj,dom,sess,maskV,maskType,targTpeak,targEcc), summarise,
                threshMean = mean(thresh))
head(thresh)
```

### Centered data set
```{r}
dsc <- ds
centCols <- c('dom','targTpeak','stairStart','targEcc','maskV')
dsc[,centCols] <- cent(dsc[,centCols])
# dsc$maskV <- (dsc$maskV / max(dsc$maskV)) #*2 - 1
dsc$maskV_c <- (dsc$maskV + 1) / 2
dsc$targEcc_c <- (dsc$targEcc + 1) / 2
dsc$targTpeak_c <- (dsc$targTpeak + 1) / 2
dsc$subj <- as.factor(dsc$subj)
head(dsc)
```

# Plotting across target eccentricities

## Foveal targets
```{r}
ss <- sumFn(thresh[thresh$maskType=='static',])
ssAve <- ss[ss$subj=='average',]
ssIndiv <- ss[ss$subj!='average',]
pStat <- themefy(plotAve(ssAve)) + ylim(yLim)
pStatNorm <- themefy(plotAve(ssAve, ylab=yLab2, yStr='norm', seStr='seNorm')) + 
    ylim(yLimNorm)
```

### Individual plots
```{r, fig.height=6, fig.width=7}
(plotIndiv(ssIndiv))
```

### Normalized individual plots
```{r, fig.height=6, fig.width=7}
(plotIndiv(ssIndiv, ylab=yLab2, yStr='norm', seStr='seNorm'))
```

## Perifoveal targets
```{r, fig.height=3, fig.width=7}
ss <- sumFn(thresh[thresh$maskType=='dynamic',])
ssAve <- ss[ss$subj=='average',]
ssIndiv <- ss[ss$subj!='average',]
pDyna <- themefy(plotAve(ssAve)) + ylim(yLim)
pDynaNorm <- themefy(plotAve(ssAve, ylab=yLab2, yStr='norm', seStr='seNorm')) +
    ylim(yLimNorm)
```

### Individual plots
```{r, fig.height=6, fig.width=7}
(plotIndiv(ssIndiv))
```

### Normalized individual plots
```{r, fig.height=6, fig.width=7}
(plotIndiv(ssIndiv, ylab=yLab2, yStr='norm', seStr='seNorm'))
```

## Group plots
```{r}
w <- .56
```

### Non-normalized
```{r, fig.height=2.4, fig.width=7}
grid.arrange(pStat + theme(legend.position='none') + ggtitle('a: static mask'),
             pDyna + ggtitle('b: dynamic mask'),
             ncol=2, widths=c((1-w)/2,w/2))
```

### Normalized
```{r, fig.height=2.4, fig.width=7}
grid.arrange(pStatNorm + theme(legend.position='none') + ggtitle('a: static mask'),
             pDynaNorm + ggtitle('b: dynamic mask'),
             ncol=2, widths=c((1-w)/2,w/2))
```

# Analyses

## GLM
```{r}
coef(summary(lmer(thresh ~ dom + stairStart + sess + maskV * targTpeak * targEcc + 
                 (1|subj), data=dsc)))
coef(summary(lmer(thresh ~ dom + stairStart + sess + maskV_c * targTpeak_c * targEcc_c + 
                 (1|subj), data=dsc)))
```

## BF: simple effects

### Mask type
```{r}
bfBase <- lmBF(thresh ~ subj + dom + stairStart + sess, data=dsc, whichRandom='subj')
bfTest <- lmBF(thresh ~ subj + dom + stairStart + sess + maskV, data=dsc, 
               whichRandom='subj')
as.vector(bfTest / bfBase)
```
### Target Eccentricity
```{r}
bfTest <- lmBF(thresh ~ subj + dom + stairStart + sess + targEcc, data=dsc, 
               whichRandom='subj')
as.vector(bfTest / bfBase)
```
### Target onset
```{r}
bfTest <- lmBF(thresh ~ subj + dom + stairStart + sess + targTpeak, data=dsc, 
               whichRandom='subj')
as.vector(bfTest / bfBase)
```

## BF: split data sets
### Target onset
#### Static mask
```{r}
bfBase <- lmBF(thresh ~ subj + dom + stairStart + sess,
               data=dsc[dsc$maskType=='static',], whichRandom='subj')
bfTest <- lmBF(thresh ~ subj + dom + stairStart + sess + targTpeak, 
               data=dsc[dsc$maskType=='static',], whichRandom='subj')
as.vector(bfTest / bfBase)
```
#### Dynamic mask
```{r}
bfBase <- lmBF(thresh ~ subj + dom + stairStart + sess,
               data=dsc[dsc$maskType=='dynamic',], whichRandom='subj')
bfTest <- lmBF(thresh ~ subj + dom + stairStart + sess + targTpeak, 
               data=dsc[dsc$maskType=='dynamic',], whichRandom='subj')
as.vector(bfTest / bfBase)
```

## BF: two-way interactions
### maskV X targEcc
```{r}
bfBase <- lmBF(thresh ~ subj + dom + stairStart + sess + maskV + targTpeak + 
                   targEcc, data=dsc, whichRandom='subj')
bfTest <- lmBF(thresh ~ subj + dom + stairStart + sess + maskV + targTpeak + 
                   targEcc + targEcc*maskV, data=dsc, whichRandom='subj')
as.vector(bfTest / bfBase)
```
### maskV X targOnset
```{r}
bfTest <- lmBF(thresh ~ subj + dom + stairStart + sess + maskV + targTpeak + 
                   targEcc + maskV*targTpeak, data=dsc, whichRandom='subj')
as.vector(bfTest / bfBase)
```
### targEcc X onset
```{r}
bfTest <- lmBF(thresh ~ subj + dom + stairStart + sess + maskV + targTpeak + 
                   targEcc + targEcc*targTpeak, data=dsc, whichRandom='subj')
as.vector(bfTest / bfBase)
```

## BF: three-way interaction
```{r}
bfBase <- lmBF(thresh ~ subj + dom + stairStart + sess + maskV + targTpeak + 
                   targEcc + targEcc*maskV + targTpeak*maskV + targTpeak*targEcc, 
               data=dsc, whichRandom='subj')
bfTest <- lmBF(thresh ~ subj + dom + stairStart + sess + maskV + targTpeak + 
                   targEcc + targEcc*maskV + targTpeak*maskV + targTpeak*targEcc +
                   targEcc*maskV*targTpeak, data=dsc, whichRandom='subj')
as.vector(bfTest / bfBase)
```