---
title: "Motion clouds rivalry: Target detection"
author: "Egor Ananyev"
date: "April 23, 2017"
output:
    html_document:
        toc: true
---

<style type="text/css">
body, td{
  font-family: Cambria;
}
code.r{
  font-family: Inconsolata;
}
</style>

# Preparing the data

## Loading the packages and data directories
```{r set-options, message=F}
options(width=100)
out=F
an=T
# The list of subjects, the order of conditions, and the thresholds are derived from Subjects.xlsx
library(xlsx)
library(ggplot2)
library(plyr)
library(reshape)
library(matrixStats)
library(gridExtra)
library(lme4)
library(lmerTest)
library(BayesFactor)
#library(splines)
db <- '/home/egor/Dropbox/' # on Linux
db <- '/Users/Egor/Dropbox/' # Windows
# db <- '~/Dropbox/' # on Mac
source(paste(db, 'Prog/R/myFunctions/pvalfn.R', sep=''))
```

## Plot variables
```{r, message=FALSE, warning=FALSE}
# theme for plotting:
alpha <- .7
themefy <- function(p) {
    p <- p + theme_bw() + 
         theme(panel.grid.minor.x=element_blank(), panel.grid.minor.y=element_blank(),
            axis.text=element_text(size=8), axis.title=element_text(size=9),
            legend.text=element_text(size=8), legend.title=element_text(size=9),
            legend.key = element_blank(), #legend.margin=margin(t=-.04, unit='in'),
            legend.background = element_rect(fill='transparent'),
            plot.title=element_text(face='bold'),
            legend.spacing.y=unit(.01,'cm'))
}
cc <- c('#e31a1c','#fdbf6f','#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99')
```

```{r}
xLab <- 'Target Peak Time (s)'
yLab <- 'Log Contrast Threshold'
yLab2 <- 'Normalized Contrast Threshold'
# colLab <- 'Mask\nVelocity'
colLab <- expression(paste('Mask\nVelocity (', degree, '/s)', sep=''))
colLab2 <- 'Mask Velocity Bandwidth'
colLabType <- 'Mask Type   '
yLim <- c(-1.5, -0.7)
yLimNorm <- c(-1.3, -0.7)
dodge <- position_dodge(width=0.0)
sumFn <- function(ss, subjStr='subj', xStr='targTpeak', grpStr='maskV'){
    sumSubj <- ddply(ss, c(subjStr, xStr, grpStr), summarise,
                     mnS=mean(threshMean), se=sd(threshMean)/sqrt(length(threshMean))) #, .drop=F)
    sumSubjMn <- ddply(ss, c(subjStr), summarise, mnStot=mean(threshMean)) # total mean across conditions per subj
    sumSubj <- merge(sumSubj, sumSubjMn)
    sumSubj$normS <- - sumSubj$mnS / sumSubj$mnStot # normalized subject mean
    sumSubj$seNorm <- NA
    # sumSubj$mnS[is.na(sumSubj$mnS)] <- 0
    sumGrp <- ddply(sumSubj, c(xStr, grpStr), summarise,
                  mn=mean(mnS), se=sd(mnS)/sqrt(length(mnS)),
                  norm=mean(normS), seNorm=sd(normS)/sqrt(length(normS)))
    sumGrp$subj <- 'average'
    sumSubj <- rename(sumSubj, c(mnS='mn',normS='norm'))
    sumComb <- rbind(sumGrp, subset(sumSubj, select=-mnStot))
    sumComb$se[is.na(sumComb$se)] <- 0
    sumComb
}
plotAve <- function(pss, subjStr='subj', xStr='targTpeak', grpStr='maskV', 
                    xlab=xLab, ylab=yLab, collab=colLab, yStr='mn', seStr='se'){
    pss$yMin <- pss[,yStr] - pss[,seStr]
    pss$yMax <- pss[,yStr] + pss[,seStr]
    p <- ggplot(pss, aes_string(x=xStr, y=yStr, colour=grpStr, group=grpStr,
                            ymin='yMin', ymax='yMax')) +
        geom_point(position=dodge, size=1, alpha=alpha) + geom_line(position=dodge, alpha=alpha) +
        scale_x_continuous(breaks=c(.5,1,1.5), labels=c('0.5','1','1.5')) +
        geom_linerange(position=dodge, show.legend=F, alpha=alpha) +
        labs(x=xlab, y=ylab, colour=collab) + #ylim(0,1) + 
        guides(colour=guide_legend(keyheight=.2, default.unit='inch'))
    p <- themefy(p)
}
plotIndiv <- function(pss, subjStr='subj', xStr='targTpeak', grpStr='maskV', 
                    xlab=xLab, ylab=yLab, collab=colLab2, yStr='mn', seStr='se'){
    pss$yMin <- pss[,yStr] - pss[,seStr]
    pss$yMax <- pss[,yStr] + pss[,seStr]
    p <- ggplot(pss, aes_string(x=xStr, y=yStr, colour=grpStr, group=grpStr,
                            ymin='yMin', ymax='yMax')) + 
        facet_wrap( ~ subj, ncol=4) +
        geom_point(position=dodge, size=1, alpha=alpha) + geom_line(position=dodge, alpha=alpha) +
        geom_linerange(position=dodge, show.legend=F, alpha=alpha) +
        scale_x_continuous(breaks=c(.5,1,1.5), labels=c('0.5','1','1.5')) +
        # scale_y_continuous(breaks=c(0,.25,.5,.75,1), labels=c('0','','0.5','','1'), limits=c(0,1)) +
        labs(x=xlab, y=ylab, colour=collab) + 
        guides(colour=guide_legend(keyheight=.3, default.unit='inch'))
    p <- themefy(p)
}
```

## Scaling function
```{r}
cent <- function(v){
    v <- apply(v,2,function(x){
        x <- x - mean(unique(x))
        x <- x / max(x)
    })
}
```

## Loading the data
```{r}
allDataDir <- paste(db,'Projects/mc/data_bv3/mcBv3_maskV',sep='')
dataDirs <- dir(allDataDir)
dataDirs <- dataDirs[grep('maskV',dataDirs)]
colsOfInt <- c('participant', 'dom', 'session', 'mcBv', 'targTpeak', 'targXoff2', 'stairStart', 
               'meanRev6')
df <- data.frame()
dfRevs <- data.frame()
dfIntn <- data.frame()
for(curDir in dataDirs){
    print(curDir)
    curDf <- read.csv(paste(allDataDir,'/',curDir,'/',curDir,'.csv',sep=''))
    curDf <- curDf[,colsOfInt]
    subjStairs <- dir(paste(allDataDir,'/',curDir,'/',sep=''))
    subjStairs <- subjStairs[grep('.tsv',subjStairs)]
    for(curStairFN in subjStairs){
        curStair <- paste(allDataDir,'/',curDir,'/',curStairFN,sep='')
        curRevs <- read.table(curStair, skip=1, nrows=1)
        curIntn <- read.table(curStair, skip=4, nrows=2)
        curInfo <- readLines(curStair)
        curInfoCols <- data.frame(subj=curDf$participant[1], dom=curDf$dom[1], 
                                  sess=curDf$session[1], stairStart=curInfo[33], 
                                  mcBv=curInfo[41], targTpeak=curInfo[43],
                                  targEcc=curInfo[37], targV=curInfo[31])
        curDfRevs <- cbind(curInfoCols, curRevs[,2:9])
        nTrials <- ncol(curIntn)-1
        curDfIntn <- curInfoCols[rep(seq_len(nrow(curInfoCols)), each=nTrials),]
        curDfIntn$trial <- seq(1,nTrials)
        curDfIntn$intn <- as.numeric(curIntn[1,2:(nTrials+1)])
        curDfIntn$resp <- as.numeric(curIntn[2,2:(nTrials+1)])
        rownames(curDfIntn) <- NULL
        dfRevs <- rbind(dfRevs, curDfRevs)
        dfIntn <- rbind(dfIntn, curDfIntn)
    }
    df <- rbind(df, curDf[,colsOfInt])
}
ds <- rename(df, c(participant='subj', session='sess', meanRev6='thresh',
                   mcBv='maskV', targXoff2='targEcc'))
ds$maskV <- round(ds$maskV * 60 / 35, 1)
ds$maskV[ds$maskV<0.05] <- 0
ds$maskType <- 'fast'
ds$maskType[ds$maskV<5] <- 'slow'
ds$maskType[ds$maskV<=0.01] <- 'static'
dfIntn$maskV <- round(as.numeric(levels(dfIntn$mcBv))[dfIntn$mcBv] * 60 / 35, 1)
dfIntn$maskV[dfIntn$maskV<0.05] <- 0
ds$targEcc <- round(ds$targEcc/35,1)
ds$targType <- 'foveal'
ds$targType[ds$targEcc>1] <- 'perifoveal'
head(ds)
```

## Staircase averages & convenience variables

### By velocity bandwidth
```{r}
thresh <- ddply(ds, .(subj,dom,sess,maskV,targTpeak,targEcc,targType), summarise,
                threshMean = mean(thresh))
thresh$maskV <- factor(thresh$maskV)
```

### By mask type
```{r}
threshType <- ddply(ds, .(subj,dom,sess,maskType,targTpeak,targEcc,targType), summarise,
                    threshMean = mean(thresh))
```

### Centered data set
```{r}
dsc <- ds
centCols <- c('dom','targTpeak','stairStart','targEcc')
dsc[,centCols] <- cent(dsc[,centCols])
dsc$maskV <- (dsc$maskV / max(dsc$maskV)) #*2 - 1
dsc$subj <- as.factor(dsc$subj)
dsc$slowC <- -1 # scaled and centered variable [-1,1]
dsc$slowC[dsc$maskType=='slow'] <- 1 #0
dsc$slowC[dsc$maskType=='fast'] <- 999
head(dsc)
```

<!-- ## Stair plots -->
<!-- ```{r, fig.height=30, fig.width=10} -->
<!-- (p <- ggplot(dfIntn[dfIntn$targTpeak==0.5,],  -->
<!--              aes(x=trial, y=intn, colour=targEcc, linetype=stairStart)) + -->
<!--     facet_wrap(subj ~ maskV, ncol=6) + geom_point() + geom_line() + ylim(-2,0)) -->
<!-- ``` -->

<!-- ```{r, fig.height=30, fig.width=10} -->
<!-- (p <- ggplot(dfIntn[as.numeric(as.character(dfIntn$targTpeak))==1.0,],  -->
<!--              aes(x=trial, y=intn, colour=targEcc, linetype=stairStart)) + -->
<!--     facet_wrap(subj ~ maskV, ncol=6) + geom_point() + geom_line() + ylim(-2,0)) -->
<!-- ``` -->

<!-- ```{r, fig.height=30, fig.width=10} -->
<!-- (p <- ggplot(dfIntn[dfIntn$targTpeak==1.5,],  -->
<!--              aes(x=trial, y=intn, colour=targEcc, linetype=stairStart)) + -->
<!--     facet_wrap(subj ~ maskV, ncol=6) + geom_point() + geom_line() + ylim(-2,0)) -->
<!-- ``` -->

# Plotting across mask bandwidths

## Foveal targets
```{r}
ss <- sumFn(thresh[thresh$targType=='foveal',])
ssAve <- ss[ss$subj=='average',]
ssIndiv <- ss[ss$subj!='average',]
pCent <- themefy(plotAve(ssAve)) + ylim(yLim)
pCentNorm <- themefy(plotAve(ssAve, ylab=yLab2, yStr='norm', seStr='seNorm')) + 
    ylim(yLimNorm)
```

### Individual plots
```{r, fig.height=6, fig.width=7}
(plotIndiv(ssIndiv))
```

## Perifoveal targets
```{r, fig.height=3, fig.width=7}
ss <- sumFn(thresh[thresh$targType=='perifoveal',])
ssAve <- ss[ss$subj=='average',]
ssIndiv <- ss[ss$subj!='average',]
pPeri <- themefy(plotAve(ssAve)) + ylim(yLim)
pPeriNorm <- themefy(plotAve(ssAve, ylab=yLab2, yStr='norm', seStr='seNorm')) +
    ylim(yLimNorm)
```

### Individual plots
```{r, fig.height=6, fig.width=7}
(plotIndiv(ssIndiv))
```

<!-- ## Group plots -->
<!-- ```{r, fig.height=2.4, fig.width=7} -->
<!-- grid.arrange(pCent + theme(legend.position='none') + ggtitle('a: foveal target'), -->
<!--              pPeri + ggtitle('b: perifoveal target'), -->
<!--              ncol=2, widths=c((1-w)/2,w/2)) -->
<!-- ``` -->

# Plotting across mask types

## Foveal targets
```{r}
ss <- sumFn(threshType[threshType$targType=='foveal',], grpStr='maskType')
ss$maskType <- ordered(ss$maskType, levels=c('static','slow','fast'))
ssAve <- ss[ss$subj=='average',]
ssIndiv <- ss[ss$subj!='average',]
pCentType <- themefy(plotAve(ssAve, grpStr='maskType', collab=colLabType)) + ylim(yLim)
pCentNormType <- themefy(plotAve(ssAve, ylab=yLab2, yStr='norm', seStr='seNorm',
                                 grpStr='maskType', collab=colLabType)) + ylim(yLimNorm)
```

## Perifoveal targets
```{r, fig.height=3, fig.width=7}
ss <- sumFn(threshType[threshType$targType=='perifoveal',], grpStr='maskType')
ss$maskType <- ordered(ss$maskType, levels=c('static','slow','fast'))
ssAve <- ss[ss$subj=='average',]
ssIndiv <- ss[ss$subj!='average',]
pPeriType <- themefy(plotAve(ssAve, grpStr='maskType', collab=colLabType)) + ylim(yLim)
pPeriNormType <- themefy(plotAve(ssAve, ylab=yLab2, yStr='norm', seStr='seNorm', 
                             grpStr='maskType', collab=colLabType)) + ylim(yLimNorm)
```

<!-- ## Group plots -->
<!-- ```{r, fig.height=2.4, fig.width=7} -->
<!-- grid.arrange(pCentType + theme(legend.position='none') + ggtitle('a: foveal target'), -->
<!--              pPeriType + ggtitle('b: perifoveal target'), -->
<!--              ncol=2, widths=c((1-w)/2,w/2)) -->
<!-- ``` -->

# All plots combined

```{r, fig.height=2.4, fig.width=3.5}
if(out){jpeg('s6.jpg', width=3.5, height=3.5, units='in', res=600)}
w <- .59
grid.arrange(pCentType + theme(legend.position='none') + 
                 ggtitle(' central\n targets\na') + labs(y='t Threshold') +
                 theme(axis.title.x=element_blank(), axis.text.x=element_blank(),
                       axis.title.y=element_text(hjust=-.3)),
             pPeriType + ggtitle(' peripheral\n targets\nb') + 
                 theme(axis.title.y=element_blank(), axis.text.y=element_blank(),
                       axis.title.x=element_blank(), axis.text.x=element_blank()),
             pCent + ggtitle('c') + labs(x='Target Pe', y='Log Contras') + 
                 theme(legend.position='none', axis.title.x=element_text(hjust=1.3),
                       axis.title.y=element_text(hjust=1.92)),
             pPeri + ggtitle('d') + labs(x='ak Time (s)') +                
                 theme(axis.title.y=element_blank(), axis.text.y=element_blank(),
                       axis.title.x=element_text(hjust=-.475)), ncol=2, nrow=2, widths=c((1-w)/2,w/2))
if(out){dev.off()}
```

## Analyses

### GLM
```{r}
if(an){
    ssc <- dsc[dsc$maskType!='fast',]
    pvalfn(lmer(thresh ~ dom + stairStart + sess + slowC * targTpeak * targEcc + 
                     (1|subj), data=ssc))
}
```

### BF

#### Mask type
```{r}
if(an){
    bfBase <- lmBF(thresh ~ subj + dom + stairStart + sess, data=ssc, whichRandom='subj')
    bfTest <- lmBF(thresh ~ subj + dom + stairStart + sess + slowC, data=ssc, 
                   whichRandom='subj')
    as.vector(bfTest / bfBase)
}
```

#### maskTypeS X targEcc
```{r}
if(an){
    bfBase <- lmBF(thresh ~ subj + dom + stairStart + sess + slowC + targTpeak + 
                       targEcc, data=ssc, whichRandom='subj')
    bfTest <- lmBF(thresh ~ subj + dom + stairStart + sess + slowC + targTpeak + 
                       targEcc + targEcc*slowC, data=ssc, whichRandom='subj')
    as.vector(bfTest / bfBase)
}
```

# Three masks (0,1,16)
```{r}
if(an){
    ss <- ds[ds$maskV==0 | ds$maskV==1 | ds$maskV==16.5,]
    ss$maskType <- factor(ss$maskType, c('static','slow','fast'))
    ss$targType <- factor(ss$targType)
    ss$targTpeak_b <- ss$targTpeak - .5
    ss$targEcc_b <- (ss$targEcc - .8) / 2.1
    ss$targTpeak_c <-  ss$targTpeak_b * 2 - 1
    ss$targEcc_c <-  ss$targEcc_b * 2 - 1
    pvalfn(lmer(thresh ~ dom + stairStart + sess + maskType * targTpeak_b * targEcc_b +
                     (1|subj), data=ss))
    pvalfn(lmer(thresh ~ dom + stairStart + sess + maskType * targTpeak_c * targEcc_c +
                     (1|subj), data=ss))
    pvalfn(lmer(thresh ~ maskType * targTpeak_c * targEcc_c + (1|subj), data=ss))
    pvalfn(lmer(thresh ~ targTpeak_c * targEcc_c + (1|subj), data=ss))
    pvalfn(lmer(thresh ~ targEcc_c + (1|subj), data=ss[ss$maskType=='static',]))
}
```

## Bayesian test

```{r}
ss$slow <- 0
ss$slow[ss$maskType=='slow'] <- 1
ss$fast <- 0
ss$fast[ss$maskType=='fast'] <- 1
unfactored <- ss # unfactored data set for later use with regressions
allVars <- c('dom','stairStart','sess','slow','fast','targTpeak_b','targEcc_b','subj')
ss[allVars] <- lapply(ss[allVars], factor)
```

### Regressions
In the following models, the denom model is the partial model, so the question that BF factor answers is 'how much is the full model preferred against partial model'?
```{r}
## Basic regression without interactions (even though the notation includes them) or random effects:
unfactored$subj <- factor(unfactored$subj)
generalTestBF(thresh ~ subj + slow + fast + targTpeak_b + targEcc_b + 
                  targTpeak_b*targEcc_b + 
                  slow*targTpeak_b + slow*targEcc_b + fast*targTpeak_b + fast*targEcc_b + 
                  slow*targTpeak_b*targEcc_b + fast*targTpeak_b*targEcc_b, 
                  data=unfactored, whichRandom='subj', neverExclude='subj', 
              whichModels='top')
mFull <- lmBF(thresh ~ maskType*targEcc_b*targTpeak_b, data=unfactored)
mPart <- lmBF(thresh ~ slow*fast*targEcc_b, data=unfactored)
generalTestBF(thresh ~ subj + fast + targTpeak_b + targEcc_b +  targTpeak_b*targEcc_b + 
              fast*targTpeak_b + fast*targEcc_b +  fast*targTpeak_b*targEcc_b, 
              data=unfactored[unfactored$slow==0,], whichRandom='subj', 
              neverExclude='subj', whichModels='top')
 generalTestBF(thresh ~ subj + slow + targTpeak_b + targEcc_b +  targTpeak_b*targEcc_b + 
              slow*targTpeak_b + slow*targEcc_b +  slow*targTpeak_b*targEcc_b, 
              data=unfactored[unfactored$fast==0,], whichRandom='subj', 
              neverExclude='subj', whichModels='top')
```